Arrows for Parallel Computations

### 2017-12-12 paper/content/parrows.tex
minor

\subsubsection{Glasgow parallel Haskell} \label{sec:parrows:multicore}
The GpH implementation of |ArrowParallel| is implemented in a straightforward manner, but a bit different compared to the variant from Section \ref{sec:GpHIntro}, by using |evalN :: [arr a b] -> arr [a] [b]| (definition in Appendix~\ref{utilfns}, think |zipWith ($)| on Arrows) combined with |withStrategy :: Strategy a -> a -> a| from GpH, where |withStrategy| is the same as |using :: a -> Strategy a -> a|, but with flipped parameters. Our |Conf a| datatype simply wraps a |Strategy a|, but could be extended in future versions of our DSL.
\begin{figure}[t]
\begin{code}
data Conf a = Conf (Strategy a)

instance (NFData b, ArrowChoice arr) =>
  ArrowParallel arr a b (Conf b) where
    parEvalN (Conf strat) fs =
        evalN fs >>>
        arr (withStrategy (parList strat))
\end{code}% $ %% formatting
\caption{The |ArrowParallel| instance for the GpH backend.}
\label{fig:ArrowParallelMulticoreConfigurable}
\end{figure}

\subsubsection{|Par| Monad}
\olcomment{introduce a newcommand for par-monad, "arrows", "parrows" and replace all mentions to them to ensure uniform typesetting \done, we write Arrows. also "Monad"? \done}
Just like for GpH we can easily lift the definition of |parEvalN| to Arrows. To start off, we define the |Strategy a| and |Conf a| type so we can have a configurable instance of ArrowParallel:
\begin{code}
type Strategy a = a -> Par (IVar a)
data Conf a = Conf (Strategy a)
\end{code}
Now we can once again define our |ArrowParallel| instance as follows: First, we convert our Arrows |[arr a b]| with |evalN (map (>>> arr strat) fs)| into an Arrow |arr [a] [(Par (IVar b))]| that yields composable computations in the |Par| monad, where |evalN| is once again of the type |evalN :: [arr a b] -> arr [a] [b]|. By combining the result of this Arrow with |arr sequenceA|, we get an Arrow |arr [a] (Par [IVar b])|. Then, in order to fetch the results of the different threads, we map over the |IVar|s inside the |Par| Monad with |arr (>>= mapM get)| - our intermediary Arrow is of type |arr [a] (Par [b])|. Finally, we execute the comptuation |Par [b]| by composing with |arr runPar| and get the final Arrow |arr [a] [b]|.
\begin{figure}[h]
\begin{code}
instance (NFData b, ArrowChoice arr) => ArrowParallel arr a b (Conf b) where
    parEvalN (Conf strat) fs = evalN (map (>>> arr strat) fs) >>>
                    arr sequenceA >>>
                    arr (>>= mapM Control.Monad.Par.get) >>>
                    arr runPar
\end{code} %$ %% formatting
\caption{ArrowParallel instance for the Par Monad backend.}
\label{fig:ArrowParallelParMonad}
\end{figure}

>>>>

\subsubsection{Glasgow parallel Haskell} \label{sec:parrows:multicore}
The GpH implementation of |ArrowParallel| is implemented in a straightforward manner in Fig.~\ref{fig:ArrowParallelMulticore}, but a bit different compared to the variant from Section \ref{sec:GpHIntro}, by using |evalN :: [arr a b] -> arr [a] [b]| (definition in Appendix~\ref{utilfns}, think |zipWith ($)| on Arrows) combined with |withStrategy :: Strategy a -> a -> a| from GpH, where |withStrategy| is the same as |using :: a -> Strategy a -> a|, but with flipped parameters. Our |Conf a| datatype simply wraps a |Strategy a|, but could be extended in future versions of our DSL.
\begin{figure}[t]
\begin{code}
data Conf a = Conf (Strategy a)

instance (NFData b, ArrowChoice arr) =>
  ArrowParallel arr a b (Conf b) where
    parEvalN (Conf strat) fs =
        evalN fs >>>
        arr (withStrategy (parList strat))
\end{code}% $ %% formatting
\caption{The |ArrowParallel| instance for the GpH backend.}
\label{fig:ArrowParallelMulticore}
\end{figure}

\subsubsection{|Par| Monad}
\olcomment{introduce a newcommand for par-monad, "arrows", "parrows" and replace all mentions to them to ensure uniform typesetting \done, we write Arrows. also "Monad"? \done}
Just like for GpH we can easily lift the definition of |parEvalN| for the |Par| Monad to Arrows in Fig.~\ref{fig:ArrowParallelParMonad}. To start off, we define the |Strategy a| and |Conf a| type so we can have a configurable instance of ArrowParallel:
\begin{code}
type Strategy a = a -> Par (IVar a)
data Conf a = Conf (Strategy a)
\end{code}
Now we can once again define our |ArrowParallel| instance as follows: First, we convert our Arrows |[arr a b]| with |evalN (map (>>> arr strat) fs)| into an Arrow |arr [a] [(Par (IVar b))]| that yields composable computations in the |Par| monad, where |evalN| is once again of the type |evalN :: [arr a b] -> arr [a] [b]|. By combining the result of this Arrow with |arr sequenceA|, we get an Arrow |arr [a] (Par [IVar b])|. Then, in order to fetch the results of the different threads, we map over the |IVar|s inside the |Par| Monad with |arr (>>= mapM get)| - our intermediary Arrow is of type |arr [a] (Par [b])|. Finally, we execute the comptuation |Par [b]| by composing with |arr runPar| and get the final Arrow |arr [a] [b]|.
\begin{figure}[h]
\begin{code}
instance (NFData b, ArrowChoice arr) => ArrowParallel arr a b (Conf b) where
    parEvalN (Conf strat) fs = evalN (map (>>> arr strat) fs) >>>
                    arr sequenceA >>>
                    arr (>>= mapM Control.Monad.Par.get) >>>
                    arr runPar
\end{code} %$ %% formatting
\caption{ArrowParallel instance for the Par Monad backend.}
\label{fig:ArrowParallelParMonad}
\end{figure}
